---
title: "A Guide to hdar package"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{A Guide to hdar package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

### Important Notice

<div style="border: 1px solid orange; padding: 10px; background-color: #fff3cd; color: #856404;">
  <strong>Notice:</strong> We are currently performing maintenance and improvements on the Backend service. You may experience intermittent slow responses or minor issues. Rest assured, our team is working hard to enhance your experience. Thank you for your patience!
</div>

# Introduction

The `hdar` R package provides seamless access to the WEkEO Harmonised Data Access (HDA) API, enabling users to efficiently query, download, and process data from the HDA platform.

# Accessing the HDA Service

To utilize the HDA service and library, you must first register for a WEkEO account. The HDA service is available at no cost to all WEkEO users. Creating an account allows you full access to our services, ensuring you can leverage the full capabilities of HDA seamlessly. Registration is straightforward and can be completed through the following link: [Register for WEkEO](https://www.wekeo.eu/register). Once your account is set up, you will be able to access the HDA services immediately.

# Setup

To start using the `hdar` package, you first need to install and load it in your R environment.

```{r, eval = FALSE}
# Install hdar from CRAN (if available) or GitHub
# install.packages("hdar")
# or
# devtools::install_github("eea/hdar@develop")

# Load the hdar package
library(hdar)
```

# Authentication

To interact with the HDA service, you need to authenticate by providing your username and password. The `Client` class allows you to pass these credentials directly and optionally save them to a configuration file for future use. If credentials are not specified as parameters, the client will read them from the `~/.hdarc` file.

## Creating and Authenticating the Client

You can create an instance of the `Client` class by passing your username and password directly. TThe initialization method has an optional parameter `save_credentials` that specifies whether the provided credentials should be saved in the `~/.hdarc` configuration file. By default, `save_credential`s is set to `FALSE`.

### Example: Pass User/Password at Initialization

Here is an example of how to authenticate by passing the user and password, and optionally saving these credentials:

```{r, eval=FALSE}
# Define your username and password
username <- "your_username"
password <- "your_password"

# Create an instance of the Client class and save credentials to a config file
# The save_credentials parameter is optional and defaults to FALSE
client <- Client$new(username, password, save_credentials = TRUE)
```

If the `save_credentials` parameter is set to `TRUE`, the credentials will be saved in the `~/.hdarc` file, making it easier to authenticate in future sessions without passing the credentials again.

### Example: Using Saved Credentials

Once the credentials are saved, you can initialize the Client class without passing the credentials. The client will read the credentials from the `~/.hdarc` file:

```{r, eval=FALSE}
# Create an instance of the Client class without passing credentials
client <- Client$new()
```

## Checking Authentication

Once the client is created, you can check if it has been authenticated properly by calling a method `token()` that verifies authentication. For example:

```{r, eval=FALSE}
# Example method to check authentication by getting the auth token
client$get_token()
print(client$token())
```

By using one of these methods, you can securely authenticate with the HDA service and start making requests.

# Finding Datasets

To interact with the HDA service, you will often need to find datasets available on WEkEO. The Client class provides a method called `datasets` that lists available datasets, optionally filtered by a text pattern.

## Basic Usage

The basic usage of the datasets method is straightforward. You can retrieve a list of all datasets available on WEkEO by calling the `datasets` method on an instance of the `Client` class.

```{r, eval=FALSE}
# Assuming 'client' is already created and authenticated
all_datasets <- client$datasets()
print(all_datasets)
```

## Filtering Datasets

You can also filter the datasets by providing a text pattern. This is useful when you are looking for datasets that match a specific keyword or phrase.

```{r, eval=FALSE}
# Assuming 'client' is already created and authenticated
pattern <- "climate"
filtered_datasets <- client$datasets(pattern)
print(filtered_datase
```

## Understanding the Results

The datasets method returns a list containing datasets and associated information. This information may include dataset names, descriptions, and other metadata.

# Creating a Query Template

To search for data in the HDA service, you need to create a query template. Manually creating a query template can be tedious as it involves reading documentation and learning about possible parameters. To simplify this process, the `generate_query_template` function is provided to automate the creation of query templates for a given dataset.

## Basic Usage

The generate_query_template function generates a template of a query for a specified dataset. This function fetches information about existing parameters, default values, etc., from the `/queryable` endpoint of the HDA service.

#### Example: Generating a Query Template

Here is an example of how to generate a query template for the dataset with the ID "EO:CRYO:DAT:HRSI:FSC":

```{r, eval = FALSE}
# Assuming 'client' is already created and authenticated
query_template <- client$generate_query_template("EO:CRYO:DAT:HRSI:FSC")
print(query_template)
```

## Understanding the Query Template

The generated query template includes fields with descriptive placeholder values where default values are not specified. These placeholders provide hints about the expected value format.

#### Example Query Template

```{json}
{
  "dataset_id": "EO:CRYO:DAT:HRSI:FSC",
  "cloudCover": "__### Value of string type with pattern: ^[0-9]*$",
  "productIdentifier": "__### Value of string type with pattern: ^.*$",
  "q": "__### Value of string type with pattern: ^.*$",
  "observed_start": "2021-01-01T00:00:00.000Z",
  "observed_end": "__### Value of string",
  "published_start": "__### Value of string",
  "published_end": "__### Value of string"
}
```

## Using the Generated Query Template

You can and should customize the generated query template to fit your specific needs. You can remove any unnecessary parameters and set values for the ones you need. If you decide not to edit the placeholder values and use the generated query template directly in the search function, the fields with placeholder values (the ones starting with \_\_###) will be stripped off before sending the query to the HDA service.

#### Example: Using the Query Template in a Search

Here is an example of how to use the query template in a search:

```{r, eval = FALSE}
# Modify the query template as needed
query_template$cloudCover <- "10"

# Perform the search using the modified query template
matches <- client$search(query_template)
print(sapply(matches$results, function(x) {
  list(
    "id" = x$id,
    "size" = x$properties$size,
    "location" = x$properties$location
  )
}))
```

# Searching and Downloading Data

To search for data in the HDA service, you can use the `search` function provided by the Client class. This function allows you to search for datasets based on a query and optionally limit the number of results. The search results can then be downloaded using the download method of the `SearchResults` class.

## Searching for Data

The `search` function takes a query and an optional limit parameter, which specifies the maximum number of results you want to retrieve. The function only searches for data and does not download it. The output of this function is an instance of the `SearchResults` class.

#### Example: Searching for Data

Here is an example of how to search for data using a query and limit the results to 5:

```{r, eval = FALSE}
# Assuming 'client' is already created and authenticated, 'query' is defined
matches <- client$search(query, 5)
print(matches$results)
```

## Downloading the Found Results

The `SearchResults` class has a public field results and a method called download that is responsible for downloading the found data. The `download()` function takes an output directory and an optional parameter to specify which results to download.

#### Example: Downloading All Results

```{r, eval = FALSE}
# Assuming 'matches' is an instance of SearchResults obtained from the search
matches$download("~/downloads")
```

#### Example: Downloading Selected Results

You can also specify which results to download by providing their indexes:

```{r, eval = FALSE}
# Assuming 'matches' is an instance of SearchResults obtained from the search
selected_indexes <- c(1, 3, 5)  # Download only the 1st, 3rd, and 5th results
matches$download("~/downloads", selected_indexes)
```
